---
title: "Assignment 02 - Data Viz, Wrangling and Advanced Regression"
author: "RunshiYang"
output:
  html_document:
    theme: spacelab
    highlight: tango
---

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
library(data.table)
library(tidyr)
library(kableExtra)
library(tidyverse)
library(ggplot2)
```

```{r}
fn <- "https://github.com/Runshi-Yang/JSC370/blob/main/assignments/a2/WHOAlcoholTotalPerCapita_2021-09-20v2.csv"
if (!file.exists("WHOAlcoholTotalPerCapita_2021-09-20v2.csv"))
  download.file(fn, destfile = "WHOAlcoholTotalPerCapita_2021-09-20v2.csv")
alcohol_cons <- data.table::fread("WHOAlcoholTotalPerCapita_2021-09-20v2.csv")

fn <- "https://github.com/Runshi-Yang/JSC370/blob/main/assignments/a2/life-expectancy-of-women-vs-life-expectancy-of-women.csv"
if (!file.exists("life-expectancy-of-women-vs-life-expectancy-of-women.csv"))
  download.file(fn, destfile = "life-expectancy-of-women-vs-life-expectancy-of-women.csv")
life_exp <- data.table::fread("life-expectancy-of-women-vs-life-expectancy-of-women.csv")
```
## Data Wrangling

1. Put the life expectancy data in “tidy” format by creating a new column “Sex”. You may want to use `pivot_longer` function from the `tidyr` package.
```{r}
colnames(life_exp)[4:6] <- c("Female", "Male", "Population")
life_exp <- life_exp %>%
  pivot_longer(cols = c("Female",
                        "Male"),
               names_to = "Sex",
               values_to = "Life_Expectancy")
```

2. Filter the alcohol consumption data to exclude rows with data for “Both sexes”.
```{r}
alcohol_cons <- alcohol_cons %>%
  filter(Sex != "Both sexes")
```

3. For conveinence, you may rename any variables which have complicated names.
```{r}
colnames(life_exp)[1] <- "Country"
colnames(alcohol_cons)[1] <- "Region_Code"
colnames(alcohol_cons)[2] <- "Region"
colnames(alcohol_cons)[3] <- "Country_Code"
colnames(alcohol_cons)[7] <- "Consumption_Numeric"
colnames(alcohol_cons)[8] <- "Consumption_Low_Est"
colnames(alcohol_cons)[9] <- "Consumption_High_Est"
colnames(alcohol_cons)[10] <- "Consumption_String"
```

4. Merge these datasets by country name and year
```{r}
data <- merge(
  life_exp, alcohol_cons, 
  by.x = c('Country', 'Year', 'Sex'), 
  by.y = c('Country', 'Year', 'Sex'),
  all.x = FALSE, 
  all.y = FALSE
)
```

```{r}
# before we proceed, check if there is any missing data
colSums(is.na(data))
```

5. Create a summary table showing the average and sd of life expectancy and alcohol consumption by year, and sex.
```{r, message=FALSE}
data_by_year_sex <- data %>% 
  group_by(Year, Sex) %>% 
  summarise(
    Avg_Life_Expectancy = mean(Life_Expectancy),
    Sd_Life_Expectancy = sd(Life_Expectancy),
    Avg_Consumption = mean(Consumption_Numeric),
    Sd_Consumption = sd(Consumption_Numeric)
  )
data_by_year_sex %>%
  kbl() %>%
  kable_styling()
```

6. Create a new categorical variable named “consumption_level” using the alcohol total per capita variable. For female and male separately, calculate the quartiles of alcohol consumption. Categorize consumption level as low (0-q1) medium (q1-q3), and high (q3+). To make sure the variable is rightly coded, create a summary table that contains the minimum total alcohol consumption, maximum alcohol consumption, and number of observations for each category.
```{r}
# Quartiles for females
female_quartiles <- data %>%
  filter(Sex == "Female") %>%
  summarize(q1 = quantile(Consumption_Numeric, 0.25),
            q3 = quantile(Consumption_Numeric, 0.75))

# Quartiles for males
male_quartiles <- data %>%
  filter(Sex == "Male") %>%
  summarize(q1 = quantile(Consumption_Numeric, 0.25),
            q3 = quantile(Consumption_Numeric, 0.75))

# Create a new categorical variable named “Consumption_Level”
data <- data %>% 
  filter(!is.na(Consumption_Numeric)) %>% 
  mutate(Consumption_Level = case_when(
    Sex == 'Female' & Consumption_Numeric <= female_quartiles$q1 ~ 'low',
    Sex == 'Female' & Consumption_Numeric > female_quartiles$q1 & Consumption_Numeric <= female_quartiles$q3 ~ 'medium',
    Sex == 'Female' & Consumption_Numeric > female_quartiles$q3 ~ 'high',
    Sex == 'Male' & Consumption_Numeric <= male_quartiles$q1 ~ 'low',
    Sex == 'Male' & Consumption_Numeric > male_quartiles$q1 & Consumption_Numeric <= male_quartiles$q3 ~ 'medium',
    Sex == 'Male' & Consumption_Numeric > male_quartiles$q3 ~ 'high'
    ))

# Create a summary table
summary_by_sex <- data %>% 
  group_by(Sex) %>% 
  summarise(
    Min_Consumption = min(Consumption_Numeric),
    Max_Consumption = max(Consumption_Numeric),
    Num_Observation = sum(!is.na(Consumption_Numeric))
  )

summary_by_sex %>%
  kbl() %>%
  kable_styling()
```

## Looking at the Data

1. Stacked histogram of alcohol consumption by sex.

2. Facet plot by year for 2000, 2010, and 2019 showing scatterplots with regression lines of life expectancy and alcohol consumption.

3. A linear model of life expectancy as a function of time, adjusted for sex. Compare the summary for Canada, and a second country of your choice.

4. A barplot of male and female life expectancy for the 10 countries with largest discrepancies in 2019.

5. A boxplot of life expectancy by alcohol consumption level and sex for the year 2019.

6. Choose a visualization to examine the association life expectancy with alcohol consumption over time.


## Advanced Regression

Construct a multiple linear regression model to examine the association between alcohol consumption and life expectancy, adjusted for time and sex. First use time as a linear predictor variable, and then fit another model where you put a cubic regression spline on time. Provide summaries of your models, plots of the linear and non-linear associations, and interpretation of the linear and non-linear associations.